Respiratory Outbreak Trends in Toronto Healthcare Institutions Across the COVID-19 Pandemic 

### 2025-05-14 paper/paper.qmd
Fixed typos

---
title: "Respiratory Outbreaks Trends in Toronto Healthcare Institutions Across the COVID-19 Pandemic "
subtitle: "The Response of Coronavirus and Other Respiratory Diseases to COVID-19 Pandemic Policies"
author: 
  - Evelyn Hughes
thanks: "Code and data are available at: [https://github.com/EveHughes/Donaldson-Paper]."
date: 05/12/2025
date-format: long

@@

## Analysis Data
### Yearly Outbreak Count {#sec-yearly_count}
To simulate, test, download, and clean the Outbreaks in Toronto Healthcare Institutions (2019-2024), the statistical programming language Python was used [@python]. Specific libraries that assisted the analysis include `numpy` [@numpy], `matplotlib` [@matplotlib], and `polars` [@polars].

@@

#plotting data & adding labels
t = np.arange(2019, 2025, 1)
KEYS = ["Total", "Respiratory", "Coronavirus", "Influenza", "Syncytial Virus", "Metapneumovirus", "Rhinovirus", "Parainfluenza"]
df = pl.read_csv("../data/02-analysis_data/yearly_disease_count.csv")
raw_data = df.to_dict()
for key in KEYS:
  plt.plot(t, raw_data[key], marker = 'o', label = key)
plt.title("Outbreak Counts")
plt.xlabel("Year")
plt.ylabel("Outbreak Count")
plt.grid()

@@

@fig-all_plot showcases that the total number of cases fluctuates with the number of respiratory cases, with number of outbreaks increasing since 2019 to 2022, and beginning to decrease afterwards.
Additionally, the number of Coronavirus cases also fluctuates with the number of respiratory cases and total cases, following the same trend described above.

@@

#plot data
t = np.arange(2019, 2025, 1)
KEYS = ["Influenza", "Syncytial Virus", "Metapneumovirus", "Rhinovirus", "Parainfluenza"]
df = pl.read_csv("../data/02-analysis_data/yearly_disease_count.csv")
raw_data = df.to_dict()
for key in KEYS:
  plt.plot(t, raw_data[key], marker = 'o', label = key)

>>>>

---
title: "Respiratory Outbreak Trends in Toronto Healthcare Institutions Across the COVID-19 Pandemic "
subtitle: "The Response of Coronavirus and Other Respiratory Diseases to COVID-19 Pandemic Policies"
author: 
  - Evelyn Hughes
thanks: "Code and data are available at: [https://github.com/EveHughes/Donaldson-Paper]."
date: 05/12/2025
date-format: long

@@

## Analysis Data
### Yearly Outbreak Count {#sec-yearly_count}
To simulate, test, download, and clean the Outbreaks in Toronto Healthcare Institutions (2019-2024), the programming language Python was used [@python]. Specific libraries that assisted the analysis include `numpy` [@numpy], `matplotlib` [@matplotlib], and `polars` [@polars].

@@

#plotting data & adding labels
t = np.arange(2019, 2025, 1)
KEYS = ["Total", "Respiratory", "Coronavirus", "Influenza", "RSV", "Metapneumovirus", "Rhinovirus", "Parainfluenza"]
df = pl.read_csv("../data/02-analysis_data/yearly_disease_count.csv")

df = df.rename({"Syncytial Virus": "RSV"})

raw_data = df.to_dict()
for key in KEYS:
  plt.plot(t, raw_data[key], marker = 'o', label = key)
plt.title("Outbreak Counts")
plt.xlabel("Year")
plt.ylabel("Outbreak Count")
plt.grid()

@@

@fig-all_plot showcases that the annual total number of outbreaks fluctuates with the annual number of respiratory outbreaks. The count of both of these has increased from 2019 to 2022, and began to decrease afterwards.
Additionally, the number of Coronavirus outbreaks also fluctuates with the number of respiratory outbreaks and total outbreaks, following the same trend described above.

@@

#plot data
t = np.arange(2019, 2025, 1)
KEYS = ["Influenza", "RSV", "Metapneumovirus", "Rhinovirus", "Parainfluenza"]
df = pl.read_csv("../data/02-analysis_data/yearly_disease_count.csv")
df = df.rename({"Syncytial Virus": "RSV"})

raw_data = df.to_dict()
for key in KEYS:
  plt.plot(t, raw_data[key], marker = 'o', label = key)
